Tail Flexibility in the Degrees of Preferential Attachment Networks

### 2025-06-01 paper.tex
Implementing changes

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  sn-basic,
]{sn-jnl}

\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

@@

\raggedbottom
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Tail Flexibility in the Degrees of Preferential Attachment Networks},
  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
  pdfkeywords={networks, extremes},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

% keywords
\keywords{networks,  extremes}

@@

These mechanistic models are generally useful when simulating networks,
but not for fitting to real data as usually only snapshots of a network
are available and not the full evolution. However, one statistic that is
usually available for snapshots of networks is the degree distribution
making it valuable to study the degrees in a network. Often, when
looking at the degree distributions of real networks it can be
attractive to use the power law when modelling as it is seemingly
observed in much of the data. It also comes with the added benefit of
offering a connection to preferential attachment since the BA model
results in a power law degree distribution \citep{Barabasi99}, perhaps
suggesting that if the power law fits well then the network may have
grown according to similar simple rules. However, attempting to explain
the networks growth using preferential attachment when the degrees
seemingly follow a power law may not be the best course of action as
\citep{krapivsky01} shows that a general preference function does not
lead to power law degrees and additionally, it is still being argued
whether or not a lot of real data actually follows a power law.
\citet{Broido_2019} provides some evidence that suggests that real world
networks rarely have degrees that follow a power law through the use of
statistical analysis of almost 1000 networks across a wide variety of
fields; \citet{Voitalov_2019}, on the other hand, disagrees and claims
that these kinds of networks are not nearly as rare and only appear so
as a result of an unrealistic expectation of power law without
deviations or noise.

@@

\citet{rudas07} derives the limiting degree distribution in terms of the
preference function \(b(\cdot)\) and its parameters, we build upon this
by finding a class of preference functions that result in realistic
tails i.e.~heavy but not as heavy as implied by the power law.

The remainder of this paper begins with a more detailed description of
the PA model alongside the theoretical results for the limiting degree
distribution using results from \citet{rudas07}, this section
(Section~\ref{sec-tail}) then goes on to analyse the tail behaviour of
this limiting distribution in terms of the preference function
\(b(\cdot)\) arguing that a heavy tailed degree distribution can only be
the result of an asymptotically linear preference function - ending with
a suggestion for a preference function that can guarantee heavy tails
while remaining flexible. Section~\ref{sec-model} continues on from the
previous section by considering the limiting distribution obtained when
using a preference function of the form stated at the end of
Section~\ref{sec-tail} providing examples of how the shape of the
distribution varies with respect to the model parameters as well as
visualisations for how the tail heaviness changes when these parameters
are varied - continued study of this distribution then also reveals a
connection to the general Pareto distribution (GPD).
Section~\ref{sec-sim} consists of a simulation study where networks are
simulated from the PA model for various parameter combinations and then
the model specified in Section~\ref{sec-model} is fitted to the
resulting degree distributions using Bayesian inference, demonstrating
that (for simulated data) the model parameters can be recovered from
only the degree distribution. Finally, Section~\ref{sec-real} fits the
model to some real data and provides posterior estimates for the
preference function assuming the networks evolved according to the PA
model.

@@

Special cases of this model include the Barabási-Albert (BA) model when
\(b(k) = k+\varepsilon\), which leads to a power-law degree distribution
with index 2 and the Uniform Attachment (UA) model where \(b(k)=c\)
leading to a degree distribution in the Gumbel maximum domain of
attraction.

Given conditions Equation~\ref{eq-condb1} and Equation~\ref{eq-condb2}
an expression for the survival of the limiting degree distribution can
be found in the case that \(m=1\); obtained by considering a branching
process that is equivalent to the growth of the network, as in
\citep{rudas07}. Theorem 1 from \citep{rudas07} states that for the tree
\(\Upsilon(t)\) at time \(t\):

\[
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
\] where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\).

@@

\[
\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
\] which by the previously stated theorem can also be written as:

\[
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
\] Additionally, using the fact that \(f(k) = \bar F(k-1) - \bar F(k)\)
the probability mass function (p.m.f.) can be shown to be:

@@

We are specifically interested in how the tail heaviness (\(\xi\) in the
IGPD) is affected by the preference function \(b\), \citep{shimura12}
introduces a quantity that will help in determining what domain of
attraction this discrete distribution belongs to, that is:

@@

We can show that in order for the degree distribution to be heavy
tailed, the preference function must be asymptotically linear
i.e.~\(\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0\). First we must
consider the following theorem:

@@

From Proposition~\ref{prp-omega} we know that in order for the limiting
degree distribution to have heavy tails, we must have that:

\[
\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = c>0.
\] Now, setting \(b_k = b(k)\) and \(a_k = k\):

@@

\[
\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0.
\] Therefore, in order for the degree distribution to be heavy tailed,
the preference function must be asymptotically linear. With this under
consideration we propose the following preference function:

@@

Some examples of what the degree distribution looks like are shown below
in Figure~\ref{fig-polylinsurv}:

@@

The survival (Equation~\ref{eq-polysurv}) can be connected to the IGPD
by using Stirling's approximation of the gamma function, where the IGPD
is a discretised variation of the generalised Pareto distribution (GPD)
usually dubbed the Integer GPD (IGPD) where
\(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) defined by the survival: \[
\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
\] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\).

Using Stirling's approximation we have that: \[
\frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y},
\] now utilising the expression from Equation~\ref{eq-polysurv}:

@@

\[
\bar F(k) 
\begin{cases}
=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
\end{cases}
\] meaning that for \(k\ge k_0\) the limiting degree distribution is
similar to
\(\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).

To assess how close of an approximation this is the theoretical
conditional survivals are shown in Figure~\ref{fig-approx_surv} in
colour and their IGPD approximations are shown in grey. The
approximation seems to hold up fairly well even for large degrees.

@@

Since \(\beta>0\), the shape parameter of the IGPD is positive and thus
the distribution is heavy tailed. Additionally the value of the shape
parameter \(\xi\) is shown in Figure~\ref{fig-polyheat} for various
parameter choices:

@@

The darker regions on the heat maps correspond to a heavier tail and the
lighter to a lighter tail, the red dashed line shows combinations of
\(\alpha\) and \(\beta\) that produce a limiting degree distribution
with the same tail heaviness as the Barabási-Albert model, \(\xi=0.5\).

For a given network with degree count vector
\(\pmb n = (n_0, n_1, \ldots, n_M)\) and maximum degree \(M\) the
likelihood is given by:

@@

where \(B(y,z)\) is the the beta function and \(l\ge0\) variable that
allows truncating the data such that the minimum degree is \(l\).

This likelihood allows for the parameters to be inferred using either a
frequentist or a Bayesian approach.

@@

Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2} demonstrate that using
this methodology it is possible to recover the model parameters fairly
well from only the final degree distribution of a simulated network.
This indicates that the method may also be able to be applied to the
degree distributions of real networks, estimating the model parameters
assuming they evolved according to the GPA scheme.

@@

Turning now to real data, the goal is to fit the model to the degree
distributions of real networks from various sources. Alongside fitting
this model to the degree distributions, we then compare the fit to that
of an mixture distribution that was used in \citet{Lee24}.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  sn-basic,
]{sn-jnl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

@@

\raggedbottom
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Tail Flexibility in the Degrees of Preferential Attachment Networks},
  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
  pdfkeywords={networks, discrete extremes, power law, preferential
attachment},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

% keywords
\keywords{networks,  discrete extremes,  power law,  preferential
attachment}

@@

These mechanistic models are generally useful when simulating networks,
but not for fitting to real data as usually only snapshots of a network
are available and not the full evolution - without this conclusions
about the growth are difficult to make. However, one statistic that is
usually available for snapshots of networks is the degree distribution
making it valuable to study the degrees in a network. Often, when
looking at the degree distributions of real networks it can be
attractive to use the power law \footnote{Power law:
  \(\bar F(k) = L(k)\cdot k^{-1/\xi}\), where \(\xi>0\) and \(L(k)\) is
  slowly varying in \(k\).} when modelling as it is seemingly observed
in much of the data. It also comes with the added benefit of offering a
connection to preferential attachment since the BA model results in a
power law degree distribution \citep{Barabasi99}, perhaps suggesting
that if the power law fits well then the network may have grown
according to similar simple rules. However, attempting to explain the
networks growth using preferential attachment when the degrees seemingly
follow a power law may not be the best course of action as
\citep{krapivsky01} shows that a general preference function does not
necessarily lead to power law degrees, showing that a preference
function of form \(b(k) = k^\alpha\) leads to light tails when
\(\alpha<1\), a degenerate network (a finite number of vertices end up
with all of the added edges) when \(\alpha>1\), and only yielding heavy
tails when \(\alpha=1\). Additionally, it is still being argued whether
or not a lot of real data actually follows a power law.
\citet{Broido_2019} provides some evidence that suggests that real world
networks rarely have degrees that follow a power law through the use of
statistical analysis of almost 1000 networks across a wide variety of
fields; \citet{Voitalov_2019}, on the other hand, disagrees and claims
that these kinds of networks are not nearly as rare and only appear so
as a result of an unrealistic expectation of power law without
deviations or noise. \citet{rudas07} derives the limiting degree
distribution in terms of the preference function \(b(\cdot)\) and its
parameters, we build upon this by finding a class of preference
functions that result in realistic tails i.e.~heavy but not as heavy as
implied by the power law.

@@

The remainder of this paper begins with a more detailed description of
the PA model alongside the theoretical results for the limiting degree
distribution using results from \citet{rudas07},
(Section~\ref{sec-tail}) then goes on to analyse the tail behaviour of
this limiting distribution in terms of the preference function
\(b(\cdot)\) arguing that a heavy tailed degree distribution can only be
the result of an asymptotically linear preference function shown by
Proposition~\ref{prp-omega} - ending with a suggestion for a preference
function that can guarantee heavy tails while remaining flexible.
Section~\ref{sec-model} builds on the results from
Section~\ref{sec-tail} and considers the limiting distribution obtained
when using a preference function of the form stated at the end of
Section~\ref{sec-tail} providing examples of how the shape of the
distribution varies with respect to the model parameters as well as
visualisations for how the tail heaviness changes when these parameters
are varied - continued study of this distribution then also reveals a
connection to the general Pareto distribution (GPD).
Section~\ref{sec-sim} consists of a simulation study where networks are
simulated from the PA model for various parameter combinations and then
the model specified in Section~\ref{sec-model} is fitted to the
resulting degree distributions using Bayesian inference, demonstrating
that (for simulated data) the model parameters can be recovered from
only the degree distribution. Finally, Section~\ref{sec-real} fits the
model to some real data and provides posterior estimates for the
preference function assuming the networks evolved according to the PA
model.

@@

Special cases of this model include the Barabási-Albert (BA) model when
\(b(k) = k+\varepsilon\), which leads to a power-law degree distribution
with \(\xi=1/2\) and the Uniform Attachment (UA) model where \(b(k)=c\)
leading to a degree distribution that is light tailed.

Given conditions \ref{eq-condb1} and \ref{eq-condb2} an expression for
the survival of the limiting degree distribution can be found in the
case that \(m=1\); obtained by considering a branching process that is
equivalent to the growth of the network, as in \citep{rudas07}. Theorem
1 from \citep{rudas07} states that for the tree \(\Upsilon(t)\) at time
\(t\):

\begin{equation}\phantomsection\label{eq-survlim}{
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
}\end{equation} where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\).

@@

\[
\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
\] which can also be written using Equation~\ref{eq-survlim} as:

\[
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
\] Additionally, using the fact that \(f(k) = \bar F(k-1) - \bar F(k)\)
the limiting degree distribution can be shown to be:

@@

We are interested in how the tail behaviour of the limiting degrees is
affected by the preference function \(b\).

\citep{shimura12} introduces a quantity that will help in determining
this for this discrete distribution, that is:

@@

@@

\begin{proposition}[]\protect\hypertarget{prp-omega2}{}\label{prp-omega2}

Consider a GPA model with preference function \(b(\cdot)\) satisfying
Proposition~\ref{prp-omega}. If
\(\lim_{k\rightarrow\infty}\frac{b(k)}{k} = c\), then the limiting
degree distribution has a heavy tail with index \(c\over\lambda^*\) and
this heavy tail may only be the result of a preference function of this
form (asymptotically linear).

\emph{Proof}

From Proposition~\ref{prp-omega}, we have that: \[
\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = c>0.
\] Now, setting \(b_k = b(k)\) and \(a_k = k\):

@@

\[
\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0\qquad \square.
\]

Considering Proposition~\ref{prp-omega2} we propose the following
preference function:

@@

Some examples of what the degree distribution looks like for various
parameters are shown below in Figure~\ref{fig-polylinsurv}:

@@

The survival function (\ref{eq-polysurv}) can be connected to a
discretised variation of the generalised Pareto distribution (GPD)
usually dubbed the Integer GPD (IGPD) with survival: \[
\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
\] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\), denoted as
\(X|X>u \sim  \mathrm {IGPD}(\xi, \sigma, u)\) where \(\xi\) is the
shape parameter controlling th tail heaviness.

By Equation~\ref{eq-polysurv} and using Stirling's approximation:

@@

\[
\bar F(k) 
\begin{cases}
=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
\end{cases}
\] meaning that for \(k\ge k_0\) the limiting degree distribution (for
large \(k_0^\alpha\)) is approximated by
\(\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).

To assess how close of an approximation this is, the theoretical
conditional survivals are shown in Figure~\ref{fig-approx_surv} in
colour and their IGPD approximations are shown in grey. The
approximation seems to hold up fairly well even for large degrees.

@@

In agreement with Proposition~\ref{prp-omega2}, \(\beta>0\), the shape
parameter of the IGPD is positive and thus the distribution is heavy
tailed. Additionally the value of the shape parameter \(\xi\) is shown
in Figure~\ref{fig-polyheat} for various parameter choices. The darker
regions on the heat maps correspond to a heavier tail and the lighter to
a lighter tail, the red dashed line shows combinations of \(\alpha\) and
\(\beta\) that produce a limiting degree distribution with the same tail
heaviness as the Barabási-Albert model, \(\xi=0.5\).

@@

Finally to be able to perform inference of the model parameters, we
consider a network with degree count vector
\(\pmb n = (n_0, n_1, \ldots, n_M)\) and maximum degree \(M\). We can
then write the likelihood of this model for this network as:

@@

where \(B(y,z)\) is the the beta function and \(l\ge0\) is a variable
that allows truncating the data such that the minimum degree is \(l\),
which will allow the model to be fitted whilst ignoring the influence of
the lower degrees (those less than \(l\)) as the model does not capture
the behaviour at the lower degrees since \citet{rudas07} only provides
results for the case of a preferential attachment tree.

@@

Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2} shows the estimations
demonstrates that using this methodology it is possible to recover the
model parameters fairly well from only the final degree distribution of
a simulated network. This indicates that the method may also be able to
be applied to the degree distributions of real networks, estimating the
model parameters assuming they evolved according to the GPA scheme.

@@

Turning now to real data, the goal is to fit the model to the degree
distributions of real networks from various sources and learn about the
mechanics of their growth. Alongside fitting this model to the degree
distributions, we then compare the fit to that of an mixture
distribution that was used in \citet{Lee24}, but this method has the
additional benefit of learning about a network's growth. The data
consists of 12 networks sourced from \href{konect.cc}{KONECT} and the
\href{https://networkrepository.com}{Network Data Repository}\citep{nr}:

\begin{itemize}
\tightlist
\item
  \texttt{as-caida20071105}: network of autonomous systems of the
  Internet connected with each other from the CAIDA project
\item
  \texttt{dimacs10-astro-ph} : co-authorship network from the
  ``astrophysics'' section (astro-ph) of arXiv
\item
  \texttt{ego-twitter}: network of twitter followers
\item
  \texttt{facebook-wosn-wall}: subset of network of Facebook wall pasts
\item
  \texttt{maayan-faa}: USA FAA (Federal Aviation Administration)
  preferred routes as recommended by the NFDC (National Flight Data
  Center)
\item
  \texttt{maayan-Stelzl}: network representing interacting pairs of
  proteins in Humans
\item
  \texttt{moreno-blogs-blogs}: network of URLs found on the first pages
  of individual blogs
\item
  \texttt{opsahl−openflights}: network containing flights between
  airports of the world.
\item
  \texttt{pajek-erdos}: co-authorship network around Paul Erdős
\item
  \texttt{reactome}: network of protein--protein interactions in Humans
\item
  \texttt{sx-mathoverflow}: interactions from the StackExchange site
  \href{https://mathoverflow.net/}{MathOverflow}
\item
  \texttt{topology}: network of connections between autonomous systems
  of the Internet
\end{itemize}
